---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = F,
  error = F,
  warning = F
)
library(terra)
library(rworldxtra)
library(tidyterra)
library(tidyverse)
```

# PupulationCoast

<!-- badges: start -->
<!-- badges: end -->

The goal of PupulationCoast is to generate shapefiles to calculate the population density in coastal areas (100 km from the coast) and in the inner areas of the coast


```{r Other, include=FALSE}
data("countriesHigh")
```


```{r WorldBuild, cache = T}
message("reading world")
World <- terra::vect(countriesHigh)
message("Creating new variable")
World$World <- "Yes"
message("Selecting only that new variable")
World <- World[,"World"]

message("Make valid")

World <- terra::makeValid(World)
message("Aggregating")
World <- terra::aggregate(World, by = "World", cores = 3)
message("Filling holes")
World <- terra::fillHoles(World) %>% 
  terra::project("+proj=moll")
message("Maxing valid again")
World <- terra::makeValid(World)
message("Selecting only that new variable")
World <- World[,"World"]
message("Making inner buffer")
World_Inner <- terra::buffer(World, width = -100000)

World_Inner <- terra::makeValid(World_Inner)

Coastal_World <- erase(World, World_Inner)
```

```{r Areas, cache = T}
CoastalArea <- terra::expanse(Coastal_World, unit = "km") %>% sum()
InnerArea <- terra::expanse(World_Inner, unit = "km") %>% sum()
WorldArea <- terra::expanse(World, unit = "km") %>% sum()
```



```{r, cache = T}
ggplot() + geom_spatvector(data = World, fill = "green") + theme_bw() + ggtitle("World")
```

The World area is `r prettyNum(WorldArea, big.mark = ",")` sq km

```{r, cache = T}
ggplot() + geom_spatvector(data = Coastal_World, fill = "green") + theme_bw() + ggtitle("Coastal")
```

The Costal area is `r prettyNum(CoastalArea, big.mark = ",")` sq km, that is  `r round(((CoastalArea/WorldArea)*100), 2)` percent of the world

```{r, cache = T}
ggplot() + geom_spatvector(data = World_Inner, fill = "green") + theme_bw() + ggtitle("Inner")
```

The Inner area is `r prettyNum(InnerArea, big.mark = ",")`, that is  `r round(((InnerArea/WorldArea)*100), 2)` percent of the workd

```{r, eval = F}
terra::writeVector(World_Inner, "WorldInner.shp")
terra::writeVector(Coastal_World, "CoastalWorld.shp")
```

